---
title: "via_and_EDA"
author: "Yuqi Miao"
date: "9/26/2019"
output: github_document
---
```{r}
library(tidyverse)
library(ggridges)
library(hexbin)
library(patchwork)
library(ggjoy)
```

```{r, cache=TRUE}
weather_df = 
  rnoaa::meteo_pull_monitors(c("USW00094728", "USC00519397", "USS0023B17S"),
                      var = c("PRCP", "TMIN", "TMAX"), 
                      date_min = "2017-01-01",
                      date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY", 
                      USC00519397 = "Waikiki_HA",
                      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10,
    month = lubridate::floor_date(date, unit = "month")) %>%
  select(name, id, everything())
skimr::skim(weather_df)
```

```{r}
ggplot(weather_df, aes(x = tmin, y = tmax))+
    geom_density2d()
## first set the canvas and dataset, then draw the map
plot_back = ggplot(weather_df,aes(x = tmin, y = tmax))
plot_back+geom_density2d(aes(color = name),alpha = 0.5)
## two different ways to show two dimensional density plot.
plot_back+geom_bin2d(aes(fill = name))
plot_back+geom_hex()
```

## basic plots

```{r}
plot_back = ggplot(weather_df,aes(x = tmax, y = tmin, color = name))

plot_back + 
  geom_point(alpha = 0.1)+
  geom_smooth()+
  facet_grid(.~name)
```

## distinguish arguements in or out aes()

```{r}
ggplot(weather_df) + geom_point(aes(x = tmin, y = tmax, color = "blue")) ## creating a variable that has value blue everywhere.
ggplot(weather_df) + geom_point(aes(x = tmin, y = tmax),color = "blue") ## set color as blue by hand.
```

## univariate plot

Sometimes we need to know the distribution of the variable 
---- using histogram, density and boxplot. 


```{r}
p = ggplot(weather_df)
```

```{r}
## histogram

p+geom_histogram(aes(x = tmin,fill = name), position = "dodge")
p+geom_histogram(aes(x = tmin, fill = name), position = "stack")
```


```{r}
## density plot
ggplot(weather_df, aes(x = tmax, fill = name)) + 
  geom_density(alpha = .4, adjust = .5, color = "blue")+ 
  ## adjust is used to adjust the smoothing degree of the curve similar to bin_width.
  geom_rug()

## to show the density plot of multiple variables

### violin plot with error bar
violin_plot = ggplot(weather_df, aes(x = name, y = tmax))+
  geom_violin(aes(fill = name),alpha = 0.5, color = "blue") +
  stat_summary(fun.y = mean,geom = "point", color = "darkblue")+
  stat_summary(fun.ymin = function(y){mean(y)-sd(y)},fun.ymax = function(y){mean(y)+sd(y)}, geom = "errorbar", color = "darkblue")
violin_plot
## rotate: ggjoy::geom_joy or ggridges::geom_density_ridges()
density_plot = 
  ggplot(weather_df, aes(x = tmin, y = name, color = name)) +
  geom_joy(scale = 0.5)
density_plot

```

## reproducible export

### save
still with problem in mac, no clue to the device problems.
```{r}
# dev.off()
# dev.set(3)
# ggsave("density plot by names", density_plot,width = 8, height = 5)
# ggsave("violin plot with error bar", violin_plot, width = 8, height = 5)
```

## embedding plots

set in the front of all chunks.

```{r}
knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = 0.6,
  out.width = "90%"
)
```


## vis 2

```{r}
library(patchwork)
```

### labels
```{r}
weather_df %>%
  ggplot(aes(x = tmin, y = tmax, color = name))+
  geom_point(alpha = 0.5)+
  labs(
    title = "Temp plot",
    x = "min T(C)",
    y = "max T(C) -- sqrt transformation"
    ) +
  scale_x_continuous(         ### manipulate the continuous variable ticks
      breaks = c(-15,0,15), ## set the position of ticks
      labels = c("-15C", "0C","15C"), ## set the labels of ticks
      limits = c(-20,20) ## set the range of the axis
    ) +
  scale_y_continuous(         ### or using scale_y_sqrt
    trans = "sqrt",        ## using this to draw transformed variables. 
    position = "right",
    breaks = seq(0,40,by = 10),
    limits = c(0,40)
  )
  
```



### color

```{r}
weather_df %>%
  ggplot(aes(x = tmin, y = tmax, color = name))+
  geom_point(alpha = 0.5)+
  labs(
    title = "Temp plot",
    x = "min T(C)",
    y = "max T(C) -- sqrt transformation"
    ) +
  scale_color_hue(            ### not always do, change merely by hand
    name = "Weather station", ### the name of color legend
    h = c(100,300)           ### control the range of color
  )
```

### virdirs package


```{r}
ggbase <-                                         ### add a baseplot to add things
  weather_df %>%
  ggplot(aes(x = tmin, y = tmax, color = name))+
  geom_point(alpha = 0.5)+
  labs(
    title = "Temp plot",
    x = "min T(C)",
    y = "max T(C) -- sqrt transformation"
    ) + 
  viridis::scale_color_viridis(
    name = "Location", 
    discrete = TRUE  ## set to discrete variables  
  )

ggbase

ggbase_2 <- 
  weather_df %>% 
  ggplot(aes(x = tmin,fill = name)) +
  geom_histogram()+
  viridis::scale_fill_viridis(
    name = "location",
    discrete = TRUE
  )
ggbase_2
```


## legend position

```{r}
ggbase +                      ## theme setting using theme_*() will overlap the other settings, custom setting needs to be the bottome.
  theme_minimal() +
  theme(legend.position = "bottom")

ggbase +
  theme(legend.position = "bottom") + ## order mattered!!
  theme_minimal()
 
## use theme(legend.position = "none") to remove the legend
```

## update initial setting 

```{r}
knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = 0.6,
  out.width = "90%"
)

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis")

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

theme_set(theme_bw() + theme(legend.position = "bottom"))

```


## more than one dataset

```{r}
central_park = 
  weather_df %>% 
  filter(name == "CentralPark_NY")

waikiki = 
  weather_df %>% 
  filter(name == "Waikiki_HA")

ggplot(data = waikiki, aes(x = date, y = tmax, color = name)) + 
  geom_point() + 
  geom_line(data = central_park)                  ### specify the geom_specific dataset
```


(brief aside about colors)
```{r}
waikiki %>%
  ggplot(aes(x = date, y  = tmax))+
  geom_point(color = "blue")
```

when creating a summarise plot, firstly create a summary table where contains all the data we used in the overlapped complete plot. Think about the constructure and goal of the complete plot before drawing.

## patch work

```{r}
tmax_tmin_p = 
  weather_df %>% 
  ggplot(aes(x = tmax, y = tmin, color = name)) + 
  geom_point(alpha = .5) +
  theme(legend.position = "none")

prcp_dens_p = 
  weather_df %>% 
  filter(prcp > 0) %>% 
  ggplot(aes(x = prcp, fill = name)) + 
  geom_density(alpha = .5) + 
  theme(legend.position = "none")

tmax_date_p = 
  weather_df %>% 
  ggplot(aes(x = date, y = tmax, color = name)) + 
  geom_point(alpha = .5) +
  geom_smooth(se = FALSE) + 
  theme(legend.position = "bottom")

(tmax_tmin_p + prcp_dens_p) / tmax_date_p  ## patchwork！

```

## data manipulation

using data manipulation to make the data structure more clear, then making plot.

```{r}
weather_df %>%
  mutate(name = forcats::fct_relevel(name, c("Waikiki_HA", "CentralPark_NY", "Waterhole_WA"))) %>% 
  ggplot(aes(x = name, y = tmax)) + 
  geom_violin(aes(fill = name), color = "blue", alpha = .5) + 
  theme(legend.position = "bottom")
```




### reorder
```{r}
weather_df %>%
  drop_na() %>% 
  mutate(name = forcats::fct_reorder(name, tmax, mean,.desc = TRUE)) %>% 
  ## use second variable to assign the plotting order
  ## third var to set the function to manipulate key var
  ggplot(aes(x = name, y = tmax)) +  ## to change the layer order, change the dataframe arrange order!
  geom_violin(aes(fill = name), color = "blue", alpha = .5) + 
  theme(legend.position = "bottom")

weather_df %>% 
  drop_na() %>% 
  group_by(name) %>% 
  summarise(median = median(tmax),mean = mean(tmax)) %>% 
  arrange(median)
```

After setting the goal of the plot, make new datasets to show the plots!

```{r}
weather_df %>%
  select(name, tmax, tmin) %>%  ## setting the color to a factor variable
  pivot_longer(  
    tmax:tmin,
    names_to = "observation", 
    values_to = "temp") %>% 
  ggplot(aes(x = temp, fill = observation)) +
  geom_density(alpha = .5) + 
  facet_grid(~name)
```

## examples

1. imports and tidies the PULSE data, and creates a plot showing BDI score across visits.

```{r}
pulse_data = haven::read_sas(data_file = "data/public_pulse_data.sas7bdat") %>% 
  janitor::clean_names() 
pulse_visit = pulse_data %>% 
  pivot_longer(
    bdi_score_bl:bdi_score_12m,
    names_to = "visit",
    names_prefix = "bdi_score_",
    values_to = "bdi",
  ) %>% 
  select(id,visit,bdi) %>% 
  mutate(
    visit = recode(visit,"bl"="00m"),
    visit = forcats::fct_relevel(visit,c("00m","01m","06m","12m"))
    )
  
ggplot(pulse_visit,aes(x = visit, y = bdi))+
  geom_boxplot()


```


2. pup-level visulization

```{r}
pup_data = read_csv(file = "data/FAS_pups.csv", col_types = "ciiiii") %>% 
  janitor::clean_names() %>% 
  mutate(sex = recode(sex, "1" = "male","2" = "female"))

litters = read_csv("data/FAS_litters.csv",col_types = "ccddiiii") %>% 
  janitor::clean_names() %>% 
  select(-pups_survive) %>% 
  separate(col = group, into = c("dose","day_of_tx"), sep = 3) %>% 
  mutate(
    dose = factor(dose),
    wt_gain = gd18_weight - gd0_weight,
    day_of_tx = as.numeric(day_of_tx)
  )
  
fas_data = left_join(pup_data, litters, by = "litter_number")
fas_data %>% 
  select(sex,dose, day_of_tx, pd_ears:pd_walk) %>% 
  pivot_longer(
    pd_ears:pd_walk,
    names_to = "outcome",
    values_to = "pn_days",
    names_prefix = "pd_"
  ) %>% 
  drop_na() %>%
  mutate(
    outcome = forcats::fct_reorder(outcome,pn_days,median),
    dose = forcats::fct_reorder(dose,pn_days,median)
  ) %>% 
  ggplot(aes(x = dose, y = pn_days)) +
  geom_violin() +
  facet_grid(day_of_tx~outcome)
  
    
  
```






## group_by
```{r}
weather_df %>% 
  group_by(name,month) %>% 
  summarise(n_obs = n())
```

```{r}
weather_df %>%
  pull(month) %>% 
  table

```



## (2x2 tables)

## General summaries

```{r}
weather_df %>%
  group_by(month) %>%
  summarize(
    mean_tmax = mean(tmax),
    mean_prec = mean(prcp, na.rm = TRUE),
    median_tmax = median(tmax),
    sd_tmax = sd(tmax))

```


## Grouped mutate
```{r}
weather_df %>%
  group_by(name) %>%
  mutate(centered_tmax = tmax - mean(tmax)) %>% 
  ggplot(aes(x = date, y = centered_tmax, color = name)) + 
    geom_point() 

```


## Window functions


















